---
title: "stat430 a4 q3"
author: "Yiming Shen 20891774"
date: "25/11/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
x1 = rep(c(-1,1), times=8)
x2 = rep(c(-1,1), times=4, each=2)
x3 = rep(c(-1,1), times=2, each=4)
x4 = rep(c(-1,1), each=8)
x5 = x1 * x2 * x3
x6 = x2 * x3 * x4
x7 = x1 * x3 * x4
y = c(52,74,84,97,117,52,128,106,146,87,85,100,120,171,132,100)
datq3 = data.frame(x1,x2,x3,x4,x5,x6,x7,y)
datq3
```
### (a)
```{r}
mod <- lm(y ~ x1 * x2 * x3 * x4, data = datq3)
summary(mod)

effects <- mod$coef * 2
effects = effects[-1]
effects

xpos <- qqnorm(effects)$x
ypos <- qqnorm(effects,cex=0.1)$y
text(x=xpos,y=ypos,labels=names(effects))
qqline(effects)

# get the PSE
s0 <- 1.5 * median( abs(effects) )
noise <- which(abs(effects) < 2.5*s0)
pse <- 1.5 * median( abs(effects)[noise])
pse

# get the merr
d = length(effects) / 3
merr <- -qt(0.025,df=d) * pse
merr

# significant factors
which( abs(effects) > merr)

```
Comments: Based on the Lenth's method, we found that the main effects of factors
x3, x4, x7 (assumed that three-factor interactions are negligible) and the 
interaction between x2 & x4; x2 & x7 (x1:x2:x3:x4 driven by aliased interaction 
involving x7) are significant. Due to the principle of heredity, 
I would like to include that the main effect of x2 is significant as well.


### (b)
Based on (a), we found that main effects of x3,x4,x7,x2:x4,x2:x7 (and x2) are 
significant, so we can build a model using only these significant effects as follows.

```{r}
mod2 <- lm(y ~ x3 + x4 + x7 + x2*x4 + x2*x7)
summary(mod2)
```

Based on mod and mod2 above, we believe that x1,x5,x6 among the 2^(7-3) 
fractional factorial design can be reasonably treated as if it weren't factors
at all. (We still think x2 is significant due to the significance of x2:x4 and x2:x7)
We can project our 2^(7-3) fractional factorial design onto a projected 2^4 full 
factorial design using x2,x3,x4,x7 as our factors.

so we can design a full factorial design experiment as follows:
1. there are 4 factors, so it is a 2^4 full design.
2. there are 2^10 sample units, and it is a balanced design, so each condition 
   will has 64 replicates.

```{r}
# The full model can be designed in R:
# mod3 <- lm(y ~ x2 * x3 * x4 * x7, data = datq3)
# summary(mod3)
```

### (c)
After factoring phase, we can use the method of steepest ascent/descent and
response surface design to locate optimal settings of the factors we identified.

For the N = 1000 samples, we need ensure every condition has adequate replicates.
Then we code factors and response surface experimentation can be used to
reach the goal of response optimization: a second order model can be fitted and 
the optimum can be discovered via finding the stationary point.








